---
title: "Exploring Outcomes"
author: "Goran Kuljanin"
date: "March 5, 2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Libraries

Load libraries to perform analyses.

```{r libraries}
### here for project directory navigation
library(here)

### tidyverse for data manipulation and plotting
library(tidyverse)

### psych for measure evaluation
library(psych)
```

## Load Data

Load data to investigate.

```{r data}
### training data
train <- read_csv(here("data", "training_data.csv"))
  
### development data
dev <- read_csv(here("data", "development_data.csv"))
```

## Examine Outcomes

Examine the three outcomes of interest in the training data: 

1. **High_Performer**
2. **Retained**
3. **Protected_Group**

```{r outcomes}
#### counts
### High_Performer by Retained by Protected_Group
## select data
train %>%
  ## filter by High_Performer
  filter(!is.na(High_Performer)) %>%
  ## count by all three outcome variables
  count(High_Performer, Retained, Protected_Group) %>%
  ## compute proportion
  mutate(prop = n/sum(n))


### descriptive statistics by Protected_Group
## outcome
describeBy(train[,c(2:7,9)], train$Protected_Group, digits = 2, mat = TRUE)

## effect size
cohen.d(train[,c(2:7,9)], train$Protected_Group)

### High_Performer by Retained by Protected_Group
## select data
train %>%
  ## count by all three outcome variables
  count(High_Performer, Overall_Rating)
```

## Split Data

Split original training data into two new datasets: training and testing. 

```{r dataSplit}
### extract only individuals with High_Performer scores
## choose data
hp_comp <- train %>%
  ## select outcome variables and personality items
  select(High_Performer, Retained, Protected_Group, PScale01_Q1:PScale13_Q5) %>%
  ## filter for High_Performer
  filter(!is.na(High_Performer))

### create separate data frames for training and testing
## random selection of individuals
ndx <- sample(nrow(hp_comp), round(0.70*nrow(hp_comp)))

## training data frame
hp_comp_train <- hp_comp[ndx, ]

## testing data frame
hp_comp_test <- hp_comp[-ndx, ]
```

## Build Features

Perform a factor analysis on the personality items to create features out of them.

```{r pers13}
### factor analysis
fa_res <- fa(hp_comp_train[,4:58], nfactors = 13)

### add factor scores to data
## training data
hp_comp_train <- hp_comp_train %>%
  ## combine scores
  bind_cols(as_tibble(fa_res$scores))

## testing data
hp_comp_test <- hp_comp_test %>%
  ## combine scores
  bind_cols(as_tibble(
    predict.psych(fa_res, data = hp_comp_test[, 4:58], 
                  old.data = hp_comp_train[,4:58])))
```

## Build Model

Logistic regression model for each outcome.

```{r}
### models
## High_Performer
hp_train_mod <- glm(High_Performer ~ MR1 + MR2 + MR3 + MR4 + 
    MR5 + MR6 + MR7 + MR8 + MR9 + MR10 + MR11 + MR12 + MR13,
    data = hp_comp_train, family = "binomial")

## Retained
retain_train_mod <- glm(Retained ~ MR1 + MR2 + MR3 + MR4 + 
    MR5 + MR6 + MR7 + MR8 + MR9 + MR10 + MR11 + MR12 + MR13,
    data = hp_comp_train, family = "binomial")

## Protected_Group
protect_train_mod <- glm(Protected_Group ~ MR1 + MR2 + MR3 + MR4 + 
    MR5 + MR6 + MR7 + MR8 + MR9 + MR10 + MR11 + MR12 + MR13,
    data = hp_comp_train, family = "binomial")
```

## Make Predictions

Make predictions on the test dataset.

```{r}
### predictions
## select test data
hp_comp_test <- hp_comp_test %>%
  ## compute variables
  mutate(High_Performer_Prob = predict(hp_train_mod, hp_comp_test, type = "response"),
         Retained_Prob = predict(retain_train_mod, hp_comp_test, type = "response"),
         Protected_Group_Prob = predict(protect_train_mod, hp_comp_test, type = "response"))
```

## Evaluate Score

Evaluate score of predictions.

```{r}
### filter for non-missing predictions
hp_comp_test_nomiss <- hp_comp_test %>%
  ## select variables
  select(High_Performer, High_Performer_Prob,
         Retained, Retained_Prob,
         Protected_Group, Protected_Group_Prob) %>%
  ## filter for non-missing
  filter(!is.na(High_Performer_Prob))

### select top-half of performers
hp_comp_test_nomiss <- hp_comp_test_nomiss %>%
  ## arrange by High_Performer_Prob
  arrange(desc(High_Performer_Prob)) %>%
  ## compute variable
  mutate(hire = if_else(1:nrow(hp_comp_test_nomiss) < round(0.5*nrow(hp_comp_test_nomiss)),
                        1, 0))

### model performance
## accuracy true High_Performer
hp_perf <- hp_comp_test_nomiss %>%
  filter(High_Performer == 1) %>%
  count(hire == 1)

## accuracy true Retained
retain_perf <- hp_comp_test_nomiss %>%
  filter(Retained == 1) %>%
  count(hire == 1)

## accuracy High_Performer and Retained
hp_retain_perf <- hp_comp_test_nomiss %>%
  filter(High_Performer == 1, Retained == 1) %>%
  count(hire == 1)

## non-protected group selection 
non_protect_perf <- hp_comp_test_nomiss %>%
  filter(Protected_Group == 0) %>%
  count(hire == 1)

## protected group selection
protect_perf <- hp_comp_test_nomiss %>%
  filter(Protected_Group == 1) %>%
  count(hire == 1)

## adverse impact ratio
air <- (protect_perf$n[2]/sum(protect_perf$n))/
  (non_protect_perf$n[2]/sum(non_protect_perf$n))

### final score
hp_perf$n[2]/sum(hp_perf$n)*25 + 
  retain_perf$n[2]/sum(retain_perf$n)*25 +
  hp_retain_perf$n[2]/sum(hp_retain_perf$n)*50 -
  abs(1 - air)*100
```

